---
title: "Barplots and Histograms Assignment Solutions"
author: "Mike LeVan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyverse)
```

# Barplots and Histograms Assignment Solutions

## Overview {.unnumbered}

In this assignment, you will be working with a small dataset and using R to create basic bar plots and histograms. This exercise will help you get familiar with data visualization techniques, which are essential for data analysis and interpretation.

## Skills Checked

1.  **Load the Dataset**: Load the provided dataset into R.
2.  **Bar Plot**: Create a bar plot to visualize the data.
3.  **Histogram**: Create a histogram to visualize the data.

## Barplot Dataset 1

You will be using a dataset that contains information about the number of students enrolled in different courses at a university. Copy the following dataset into a CSV file named `courses.csv` and upload it to your working directory.

| Course           | Number of Students |
|:-----------------|:------------------:|
| Mathematics      |         45         |
| Physics          |         30         |
| Chemistry        |         25         |
| Biology          |         50         |
| Computer_Science |         40         |
| History          |         35         |
| English          |         20         |
| Economics        |         30         |

## Instructions

1.  **Load the Dataset**
    -   Make sure you have the `readr` package installed.
    -   Use the `read_csv` function to load the dataset from the provided CSV file.
    -   Store the dataset in a variable called `courses_data`.

```{r message=FALSE, warning=FALSE}
   # Load the dataset
   courses_data <- read_csv("./courses.csv")
```

2.  **Create a Bar Plot**
    -   Use the `ggplot` function to create a bar plot of the number of students enrolled in each course.
    -   Label the x-axis as "Courses" and the y-axis as "Number of Students".
    -   Add a title to the plot: "Number of Students Enrolled in Each Course".

```{r}
#| code-fold: true
   # Create a bar plot
# Create a bar plot using ggplot2
ggplot(courses_data, aes(x = Course, y = Students)) +
  geom_bar(stat = "identity", fill = "blue") +
  xlab("Courses") +
  ylab("Number of Students") +
  ggtitle("Number of Students Enrolled in Each Course") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Submission

Submit the R script file containing your code for loading the dataset, creating the bar plot, and creating the histogram. Ensure that your code is well-documented with comments explaining each step.

## Evaluation

Your assignment will be evaluated based on the following criteria: - Correctness of the code - Proper use of R functions for data visualization - Clear and accurate labeling of the plots - Overall presentation and readability of the code

Good luck!
