
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(readxl)
```

# Barplots and Histograms Assignment   {.unnumbered}

In this assignment, you will be working with a small dataset and using R to create basic bar plots and histograms. This exercise will help you get familiar with data visualization techniques, which are essential for data analysis and interpretation.

## Skills Checked

1.  **Load the Dataset**: Load the provided dataset into R.
2.  **Bar Plot**: Create a bar plot to visualize the data.
3.  **Histogram**: Create a histogram to visualize the data.

## Barplot Dataset 1

The following is a dataset that contains information about the number of 
students enrolled in different courses at a university. Copy the following 
dataset into a CSV file named `courses.csv` and upload it to your working 
directory. Create an appropriate barchart depicting the data.

| Course           | Number of Students |
|:-----------------|:------------------:|
| Mathematics      |         45         |
| Physics          |         30         |
| Chemistry        |         25         |
| Biology          |         50         |
| Computer_Science |         40         |
| History          |         35         |
| English          |         20         |
| Economics        |         30         |

## Barplot Dataset 2

The following is a dataset that contains sales information about the number of 
electronic sales at your company. Copy the following 
dataset into a CSV file named `sales.csv` and upload it to your working 
directory.  Create an appropriate barchart depicting the data.

| Product           | Sales |
|:-----------------|:------------------:|
| Laptops      |         120         |
| Smartphones          |         200         |
| Tablets      |         80         |
| Accessories          |         150         |
| Wearables |         90         |
| Desktops        |         70         |
| Cameras         |         60         |
| Printers        |         40         |

## Barplot Dataset 3

The following is a dataset that contains 
monthly rainfall (in millimeters) information for various
cities across the United States. Copy the following 
dataset into a CSV file named `rainfall`.csv` and upload it to your working 
directory.  Create an appropriate barchart depicting the data.

| City           | Monthly_Rainfall_mm |
|:-----------------|:------------------:|
| New_York  |         120         |
| Los_Angeles   |         20         |
| Chicago |         80         |
| Houston |         90         |
| Phoenix |         15         |
| Philadelphia    |         100         |
| San_Antonio        |         70         |
| San_Diego      |         30         |
| Dallas | 85 |
| San_Jose | 25 |

## Barplot Dataset 4

The following is a dataset that contains 
monthly transatlantic airtravel, in thousands of passengers, for 1958-1960. 
There are 4 fields, "Month", "1958", "1959" and "1960" and 12 records, 
"JAN" through "DEC". 
<br>

Create a bar chart that shows the average number of passengers for each month across the years 1958-1960.
<br>

Data : 
https://people.sc.fsu.edu/~jburkardt/data/csv/airtravel.csv
<br>

You will need to downlaod the data set and load it into R.


## Barplot Dataset 5

Go to Kaggle and download the dat set "US Christmas Tree Sales Data" from the 
following link:
<br>

https://www.kaggle.com/datasets/thedevastator/us-christmas-tree-sales-data/data
<br>

Create a bar chart that shows the number of trees sold (per million) for each 
year from 2010-2016.


```{r message=FALSE, warning=FALSE}
tree_data <- read_csv("./US_Christmas_Tree_Sales_2010_to_2016.csv")
```

## Histogram Dataset 1

The following is a dataset (`Ages.csv`) that contains information about the ages of 
participants at a local community center. Copy the dataset into a CSV 
file named `ages_data.csv` and upload it to your working directory. 
Create a  histogram with 7 bins depicting the data.

```{r message=FALSE, warning=FALSE}
ages_data <- read_csv("./Ages.csv")

ages_data
```

## Histogram Dataset 2

The following  (`Lexington_Temperature_Data.csv`)
is a dataset that contains information about the 
daily temperature in Celsius for Lexington the first 50
days of spring.
Copy the dataset into a CSV 
file named `Lex_temps.csv` and upload it to your working directory. 
Create a histogram with 4 bins depicting the data.

```{r message=FALSE, warning=FALSE}
Lex_temps <- read_csv("./Lexington_Temperature_Data.csv")

Lex_temps
```


## Histogram Dataset 3

The following is a data set (`kentucky_mens_basketball_wins.csv`) that contains information about the
number of wins for the University of Kentucky's men's basketball team from 1980 - 2024.
Copy the dataset into a CSV
file named `UK_wins` and upload it to your working directory.
Create a histogram with 10 bins depicting the data on the amount of wins
per year.

```{r message=FALSE, warning=FALSE}

UK_wins <- read_csv("./kentucky_mens_basketball_wins.csv")

UK_wins
```

## Histogram Dataset 4

The following is a data set (`Fire_Arm_Deaths_1990_to_2022.xlsx`) that contains information about the
number of firearm deaths in the US from 1990 - 2024.
Copy the dataset into a CSV
file named `Fire_Arm_Deaths` and upload it to your working directory.
Create a histogram with 10 bins depicting the data on the amount of deaths 
per year.

```{r message=FALSE, warning=FALSE}
library(readxl)

Fire_Arm_Deaths <- read_csv("./Fire_Arm_Deaths_1990_to_2022.csv")

Fire_Arm_Deaths
```


## Histogram Dataset 5

The data set (`gdp-per-capita-us-dollars-2020.csv`) contains 
information about the GDP per capita in US dollars for various countries in 2020.
Download the data set from 
<a href="https://www.datasetsanalysis.com/descriptive-statistics/real-life-data/gdp-per-capita-us-dollars-2020.csv">here</a>
and save it to yourworking directory.
Copy the dataset into a variable named `gdp_per_capita`.

```{r message=FALSE, warning=FALSE}

gdp_per_capita <- read_csv("./gdp-per-capita-us-dollars-2020.csv")

head(gdp_per_capita)
```

  - Create a histogram, with equal class widths, with the first class being (0 , 10000 ] and covers all data values.
  - Create a histogram, with equal class widths, with the first class being (0 , 5000] with the last class being (55000 , 60000]
  - Create a histogram, with equal class widths, with the first class being (0 , 1000] with the last class being (19000 , 20000]
  - Compare all three histograms. Which one gives more information as to how the countries with low GDP are distributed?



