---
title: "Beginning Data Cleaning"
author: "Mike LeVan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Beginning Data Cleaning

What is "data cleaning"? Data cleaning is the process of preparing data 
for analysis. This process includes:

1. Removing missing values
2. Removing duplicates
3. Removing outliers
4. Standardizing data
5. Normalizing data
6. Encoding categorical variables
7. Creating new variables
8. Merging datasets
9. Reshaping datasets
10. Splitting datasets

In this tutorial, we will focus on the first three steps: removing missing 
values, removing duplicates, and removing outliers.

### Removing Missing Values

Missing values are values that are not present in the dataset. Missing 
values can be represented in many ways, including:

- `NA`
- `NaN`
- `NULL`
- `""`
- `0`
- `-1`
- `9999`

To remove missing values from a dataset, we can use the `na.omit()` 
function in R. This function removes rows with missing values from a dataset.

```{r}

# Create a dataset with missing values

data <- data.frame(
  x = c(1, 2, NA, 4, 5),
  y = c(NA, 2, 3, 4, 5)
)

# Remove missing values

data_clean <- na.omit(data)

# Print the cleaned dataset

print(data_clean)

```

### Removing Duplicates

Duplicates are rows that are identical in a dataset. To remove duplicates 
from a dataset, we can use the `unique()` function in R. This function 
removes duplicate rows from a dataset.

```{r}

# Create a dataset with duplicates

data <- data.frame(
  x = c(1, 2, 3, 4, 5, 1, 2, 3, 4, 5),
  y = c(1, 2, 3, 4, 5, 1, 2, 3, 4, 5)
)

# Remove duplicates

data_clean <- unique(data)

# Print the cleaned dataset

print(data_clean)

```

### Removing Outliers

Outliers are values that are significantly different from the rest of 
the data in a dataset. To remove outliers from a dataset, we can use 
the `boxplot()` function in R. This function creates a boxplot of the 
data, which can help us identify outliers.

```{r}

# Create a dataset with outliers

data <- data.frame(
  x = c(1, 2, 3, 4, 5, 100),
  y = c(1, 2, 3, 4, 5, 100)
)

# Create a boxplot of the data

boxplot(data)

```

From the boxplot, we can see that the value `100` is an outlier. To remove 
this outlier, we can use the `subset()` function in R. This function 
removes rows that meet a certain condition from a dataset.

```{r}

# Remove outliers

data_clean <- subset(data, x < 100 & y < 100)

# Print the cleaned dataset

print(data_clean)

```

In this tutorial, we learned how to remove missing values, duplicates, 
and outliers from a dataset. In the next tutorial, we will focus on 
standardizing and normalizing data.

---

## References

- Wickham, H. (2014). Tidy data. Journal of Statistical Software, 59(10), 1-23.

- Wickham, H. (2019). R for Data Science. O'Reilly Media.

- Grolemund, G., & Wickham, H. (2017). R for Data Science. O'Reilly Media.

- Kuhn, M., & Johnson, K. (2013). Applied Predictive Modeling. Springer.

- James, G., Witten, D., Hastie, T., & Tibshirani, R. (2013). An Introduction to Statistical Learning. Springer.

- Hastie, T., Tibshirani, R., & Friedman, J. (2009). The Elements of Statistical Learning. Springer.

- Zuur, A., Ieno, E. N., & Smith, G. M. (2007). Analysing Ecological Data. Springer.

- Venables, W. N., & Ripley, B. D. (2002). Modern Applied Statistics with S. Springer.


```

