---
title: "dplyr and Piping Assignment"
author: "Mike LeVan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Exercises

**Instructions:** Solve each of the following problems using the commands 
we saw earlier from the dplyr package : 

-   chaining operations with the `%>%` (piping) operator
-   `filter()`: Filter rows based on conditions
-   `mutate()`: Create new columns
-   `select()`: Select columns
-   `arrange()`: Arrange rows
-   `group_by()`: Group data
-   `summarize()`: Summarize data

Each of the solutions should be one piped command. Some may be a single
pipe  while others may take multiple pipes. 

### Problem 1: Filtering Rows
Use the `mtcars` dataset to filter the rows where the `mpg` (miles per gallon) is greater than 20.
Save the result to `filtered_mtcars`.

```{r}
# Solution

# If needed, load up dplyr

# library(dplyr)

filtered_mtcars <- mtcars %>%
  filter(mpg > 20)

filtered_mtcars
```

### Problem 2: Selecting Columns
Use the `mtcars` dataset to select the columns `mpg`, `cyl`, and `hp`. Save the 
result to `selected_mtcars`.

```{r}

# Solution

selected_mtcars <- mtcars %>%
  select(mpg, cyl, hp)

selected_mtcars
```

### Problem 3: Arranging Rows
Use the `mtcars` dataset to arrange the rows in descending order of `hp` (horsepower).
Save the result to `arranged_mtcars`.

```{r}

# Solution

arranged_mtcars <- mtcars %>%
  arrange(desc(hp))

arranged_mtcars
```

### Problem 4: Creating New Columns
Use the `mtcars` dataset to create a new column `power_to_weight` which is the ratio of `hp` (horsepower) to `wt` (weight).
Save the result to `mutated_mtcars`.

```{r}

# Solution

mutated_mtcars <- mtcars %>%
  mutate(power_to_weight = hp / wt)

mutated_mtcars
```

### Problem 5: Summarizing Data
Use the `mtcars` dataset to calculate the mean `mpg` (miles per gallon).
Save the result to `summary_mtcars`.

```{r}

# Solution

summary_mtcars <- mtcars %>%
  summarise(mean_mpg = mean(mpg))

summary_mtcars
```

### Problem 6: Grouping and Summarizing Data
Use the `mtcars` dataset to calculate the mean `mpg` (miles per gallon) for each number of cylinders (`cyl`).
Save the result to `grouped_summary_mtcars`.

```{r}

# Solution

grouped_summary_mtcars <- mtcars %>%
  group_by(cyl) %>%
  summarise(mean_mpg = mean(mpg))

grouped_summary_mtcars
```

### Problem 7: Filtering, Selecting, and Arranging
Use the `mtcars` dataset to filter the rows where `mpg` is greater than 20, select the columns `mpg` and `hp`, and arrange the rows in ascending order of `hp`. Save the result to `filtered_selected_arranged_mtcars`.

```{r}
# Solution

filtered_selected_arranged_mtcars <- mtcars %>%
  filter(mpg > 20) %>%
  select(mpg, hp) %>%
  arrange(hp)

filtered_selected_arranged_mtcars
```

### Problem 8: Creating Multiple New Columns
Use the `mtcars` dataset to create two new columns: `power_to_weight` (ratio of `hp` to `wt`) and `mpg_per_cyl` (ratio of `mpg` to `cyl`).
Save the result to `mutated_mtcars`.

```{r}
# Solution

mutated_mtcars <- mtcars %>%
  mutate(power_to_weight = hp / wt,
         mpg_per_cyl = mpg / cyl)

mutated_mtcars
```

### Problem 9: Counting Rows
Use the `mtcars` dataset to count the number of cars for each number of cylinders (`cyl`). 
Save the result to `counted_mtcars`.

```{r}

# Solution

counted_mtcars <- mtcars %>%
  count(cyl)

counted_mtcars
```

### Problem 10: Filtering and Summarizing
Use the `airquality` dataset to filter the rows where `Month` is 5 (May), and calculate the mean `Temp` (temperature).
Save the result to `filtered_summary_airquality`.


```{r}

# Solution

data("airquality")

filtered_summary_airquality <- airquality %>%
  filter(Month == 5) %>%
  summarise(mean_temp = mean(Temp, na.rm = TRUE))

filtered_summary_airquality
```

### Problem 11: Grouping and Creating New Columns
Use the `airquality` dataset to group the data by `Month` and create a new column `mean_temp` which is the mean `Temp` (temperature) for each month. Save the result to `grouped_mutated_airquality`. 

```{r}

# Solution

grouped_mutated_airquality <- airquality %>%
  group_by(Month) %>%
  mutate(mean_temp = mean(Temp, na.rm = TRUE))

grouped_mutated_airquality
```

### Problem 12: Filtering, Grouping, and Summarizing
Use the `airquality` dataset to filter the rows where `Temp` is greater than 80, group by `Month`, and calculate the mean `Ozone` level for each month. Save the result to `filtered_grouped_summary_airquality`.

```{r}

# Solution

filtered_grouped_summary_airquality <- airquality %>%
  filter(Temp > 80) %>%
  group_by(Month) %>%
  summarise(mean_ozone = mean(Ozone, na.rm = TRUE))

filtered_grouped_summary_airquality
```

### Problem 13: Selecting and Summarizing
Use the `trees` dataset to select the `Height` and `Volume` columns and calculate the mean `Height` and `Volume`.
Save the result to `selected_summary_trees`.

```{r}

# Solution

data("trees")

selected_summary_trees <- trees %>%
  select(Height, Volume) %>%
  summarise(mean_height = mean(Height),
            mean_volume = mean(Volume))

selected_summary_trees
```

### Problem 14: Filtering, Mutating, and Arranging
Use the `trees` dataset to filter the rows where `Height` is greater than 75, create a new column `volume_to_height` which is the ratio of `Volume` to `Height`, and arrange the rows in descending order of `volume_to_height`. Save the result to `filtered_mutated_arranged_trees`.

```{r}
# Solution

filtered_mutated_arranged_trees <- trees %>%
  filter(Height > 75) %>%
  mutate(volume_to_height = Volume / Height) %>%
  arrange(desc(volume_to_height))

filtered_mutated_arranged_trees
```

### Problem 15: Combining Multiple Steps
Use the `trees` dataset to filter the rows where `Height` is greater than 70, select the `Height` and `Volume` columns, create a new column `volume_to_height` which is the ratio of `Volume` to `Height`, and calculate the mean `volume_to_height`.
Save the result to `combined_trees`.

```{r}

# Solution

combined_trees <- trees %>%
  filter(Height > 70) %>%
  select(Height, Volume) %>%
  mutate(volume_to_height = Volume / Height) %>%
  summarise(mean_volume_to_height = mean(volume_to_height))
combined_trees

```