---
title: "Intermediate Data Cleaning"
author: "Mike LeVan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Intermediate Data Cleaning

In this tutorial, we will learn how to clean data using the `dplyr` 
package. We will use the `gapminder` dataset, which is included in the 
`gapminder` package.

```{r}

# Load the gapminder package

library(gapminder)

# Load the dplyr package

library(dplyr)

# View the first few rows of the gapminder dataset

head(gapminder)

```

### Selecting Columns

```{r}

# Select the country and year columns

gapminder %>%
  select(country, year)

```

### Filtering Rows

```{r}

# Filter the dataset to include only rows where the country is "United States"

gapminder %>%
  filter(country == "United States")

```

### Creating New Columns

```{r}

# Create a new column called "gdp" that is equal to "gdpPercap" multiplied by "pop"

gapminder %>%
  mutate(gdp = gdpPercap * pop)

```

### Grouping Data

```{r}

# Group the data by "year" and calculate the average "gdpPercap" for each year

gapminder %>%
  group_by(year) %>%
  summarise(avg_gdpPercap = mean(gdpPercap))

```

### Sorting Data

```{r}

# Sort the data by "year" in descending order

gapminder %>%
  arrange(desc(year))

```

### Combining Operations

```{r}

# Filter the data to include only rows where the country is "United States",

# group the data by "year", and calculate the average "gdpPercap" for each year

gapminder %>%
  filter(country == "United States") %>%
  group_by(year) %>%
  summarise(avg_gdpPercap = mean(gdpPercap))

```

### Conclusion

In this tutorial, we learned how to clean data using the `dplyr` 
package. We covered selecting columns, filtering rows, creating 
new columns, grouping data, sorting data, and combining operations. 
These are just a few of the many functions available in the `dplyr` 
package for data cleaning and manipulation. For more information, 
check out the `dplyr` documentation at https://dplyr.tidyverse.org/.
```

