
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises 

This exercise set will review the basic functions of R. Most of the commands 
were reviewed in the lesson, but there are a few that you will have to look up
to see how to carry out the commands.
<br>

**Instructions**

- Write R code to solve each of the following problems.
- Make sure to test your code to ensure it works correctly.
- Provide comments in your code to explain your logic where necessary.

Problem 1: Addition
Add two numbers, 12 and 34.

```{r}

# Response

```
<br>

Problem 2: Subtraction

Subtract 25 from 75.

```{r}

# Response

```
<br>

 Problem 3: Multiplication
 
Multiply 28 by 19.

```{r}

# Response

```
<br>

 Problem 4: Division
 
Divide 144 by 12.

```{r}

# Response

```
<br>

 Problem 5: Exponentiation
 
Calculate 9 raised to the power of 7.

```{r}

# Response

```
<br>

 Problem 6: Square Root
 
Find the square root of 64.

```{r}

# Response

```
<br>

 Problem 7: Logarithm
 
Calculate the natural logarithm (base e) of 20.

```{r}

# Response

```
<br>

 Problem 8: Absolute Value
 
Find the absolute value of -15.

```{r}

# Response

```
<br>

 Problem 9: Factorial
 
Calculate the factorial of 6.

```{r}

# Response

```
<br>

 Problem 10: Variables
 
Create a variable `x` that is assigned the value 10 and another variable `y` 
that is assigned the value 5. Add the two variables together.

```{r}

# Response

```
<br>

 Problem 11: Variables
 
Create a variable `z` that is the combination of `11*x + 13*y`, where `x` 
and `y` are from the previous problem.

```{r}

# Response

```
<br>

 Problem 12: Mean and Median of a Vector
 
Find the mean and median of the elements in the vector c(2, 4, 6, 8, 10, 12, 14, 16, 18, 20).

```{r}

# Response

```
<br>

Problem 13: Create a Vector

Create a vector (with an appropriate name) that contains the names of 5 of your friends. What is the
class of this vector? 

```{r}

# Response

```
<br>

 Problem 14:
 
 Create a vector with 20 random values between 300 and 500. What is the mean of this vector?
 What is the median of this vector?
 
```{r}

# Response

```
<br>

 Problem 15: Maximum and Minimum
 
Find the maximum and minimum values of the vector created in Problem 14.

```{r}

# Response

```
<br> 

Problem 16: Vectors
 
In the vector you created in problem 14 : 

  - Print out the 7th element.
  - Print out the first 10 elements.
  - Print out the last 5 elements.
  - Add the 4th and 18th elements together.

  
```{r}

# Response

```
<br>

 Problem 17: Loading Libraries
 
Load the library `readr` and use the help file to find out what 
the `read_csv` function does.

```{r}

# Response

```
<br>

 Problem 18: Boolean Vectors

Create a boolean vector that contains the values TRUE, FALSE, TRUE, TRUE, FALSE.
Print out the vector.

```{r}

# Response

```
<br>

Problem 19: Logical Operators
 
```{r}

# Explain the output for the following code :

x <- TRUE

y <- FALSE

z <- TRUE

x & (y | z)
 
```
<br>

Problem 20: Logical Operators
 
```{r}
 
# Explain the output for the following code :

x <- 5

y <- 10

z <- 15

x > y & y < z
 
```
 
 <br>
 

 
