
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercises

In this assignment, you will practice reading in data, calculating descriptive statistics (mean and median), handling missing values (NA), and using the `mutate` function to create new data columns. You will use built-in datasets from R for this assignment.

### Problem 1: Mean and Median of Sepal Length in the `iris` Dataset

**Task:** Calculate the mean and median of the `Sepal.Length` column in the `iris` dataset.
<br>

**Steps:**

1. Load the `iris` dataset.
2. Calculate the mean and median of the `Sepal.Length` column.

**Code Example:**
```{r}
# Load iris dataset
data(iris)

# Calculate mean and median of Sepal.Length
mean_sepal_length <- mean(iris$Sepal.Length)
median_sepal_length <- median(iris$Sepal.Length)

mean_sepal_length
median_sepal_length
```

### Problem 2: Mean and Median of Ozone Levels in the `airquality` Dataset

**Task:** Calculate the mean and median of the `Ozone` column in the `airquality` dataset after removing NA values.
<br>

**Steps:**

1. Load the `airquality` dataset.
2. Remove NA values from the `Ozone` column.
3. Calculate the mean and median of the `Ozone` column.

**Code Example:**
<br>
```{r}
# Load airquality dataset
data(airquality)

# Remove NA values from Ozone column
ozone_no_na <- na.omit(airquality$Ozone)

# Calculate mean and median of Ozone
mean_ozone <- mean(ozone_no_na)
median_ozone <- median(ozone_no_na)

mean_ozone
median_ozone
```

### Problem 3: Mean and Median of Annual Lynx Trappings in the `lynx` Dataset

**Task:** Calculate the mean and median of the annual number of lynx trapped in the `lynx` dataset.
<br>

**Steps:**

1. Load the `lynx` dataset.
2. Calculate the mean and median of the lynx trappings.

**Code Example:**

```{r}
# Load lynx dataset
data(lynx)

# Calculate mean and median of lynx trappings
mean_lynx <- mean(lynx)
median_lynx <- median(lynx)

mean_lynx
median_lynx
```

### Problem 4: Mean and Median of Tooth Length in the `ToothGrowth` Dataset

**Task:** Calculate the mean and median of the `len` column in the `ToothGrowth` dataset.
<br>

**Steps:**

1. Load the `ToothGrowth` dataset.
2. Calculate the mean and median of the `len` column.

**Code Example:**

```{r}
# Load ToothGrowth dataset
data(ToothGrowth)

# Calculate mean and median of len
mean_len <- mean(ToothGrowth$len)
median_len <- median(ToothGrowth$len)

mean_len
median_len
```

### Problem 5: Mean and Median of Age in the `infert` Dataset

**Task:** Calculate the mean and median of the `age` column in the `infert` dataset.
<br>

**Steps:**

1. Load the `infert` dataset.
2. Calculate the mean and median of the `age` column.

**Code Example:**

```{r}
# Load infert dataset
data(infert)

# Calculate mean and median of age
mean_age <- mean(infert$age)
median_age <- median(infert$age)

mean_age
median_age
```

## Problem 6: Mean and Median of Wind Speed in the `airquality` Dataset

**Task:** Calculate the mean and median of the `Wind` column in the `airquality` dataset after removing NA values.
<br>

**Steps:**

1. Load the `airquality` dataset.
2. Remove NA values from the `Wind` column.
3. Calculate the mean and median of the `Wind` column.

**Code Example:**

```{r}
# Load airquality dataset
data(airquality)

# Remove NA values from Wind column
wind_no_na <- na.omit(airquality$Wind)

# Calculate mean and median of Wind
mean_wind <- mean(wind_no_na)
median_wind <- median(wind_no_na)

mean_wind
median_wind
```

### Problem 7: Create a New Column and Calculate Mean and Median in the `iris` Dataset

**Task:** Create a new column `Sepal.Area` as the product of `Sepal.Length` and `Sepal.Width` in the `iris` dataset. Calculate the mean and median of the `Sepal.Area` column.
<br>

**Steps:**

1. Load the `iris` dataset.
2. Create a new column `Sepal.Area`.
3. Calculate the mean and median of the `Sepal.Area` column.

**Code Example:**

```{r}
# Load iris dataset
library(dplyr)
data(iris)

# Create new column Sepal.Area
iris <- mutate(iris, Sepal.Area = Sepal.Length * Sepal.Width)

# Calculate mean and median of Sepal.Area
mean_sepal_area <- mean(iris$Sepal.Area)
median_sepal_area <- median(iris$Sepal.Area)

mean_sepal_area
median_sepal_area
```

### Problem 8: Mean and Median of Temperature in the `airquality` Dataset

**Task:** Calculate the mean and median of the `Temp` column in the `airquality` dataset after removing NA values.
<br>

**Steps:**

1. Load the `airquality` dataset.
2. Remove NA values from the `Temp` column.
3. Calculate the mean and median of the `Temp` column.

**Code Example:**

```{r}
# Load airquality dataset
data(airquality)

# Remove NA values from Temp column
temp_no_na <- na.omit(airquality$Temp)

# Calculate mean and median of Temp
mean_temp <- mean(temp_no_na)
median_temp <- median(temp_no_na)

mean_temp
median_temp
```

### Problem 9: Create a New Column and Calculate Mean and Median in the `ToothGrowth` Dataset

**Task:** Create a new column `Dose.Milligrams` as the product of `dose` and `len` in the `ToothGrowth` dataset. Calculate the mean and median of the `Dose.Milligrams` column.
<br>

**Steps:**

1. Load the `ToothGrowth` dataset.
2. Create a new column `Dose.Milligrams`.
3. Calculate the mean and median of the `Dose.Milligrams` column.

**Code Example:**

```{r}
# Load ToothGrowth dataset
library(dplyr)
data(ToothGrowth)

# Create new column Dose.Milligrams
ToothGrowth <- mutate(ToothGrowth, Dose.Milligrams = dose * len)

# Calculate mean and median of Dose.Milligrams
mean_dose_mg <- mean(ToothGrowth$Dose.Milligrams)
median_dose_mg <- median(ToothGrowth$Dose.Milligrams)

mean_dose_mg
median_dose_mg
```

### Problem 10: Mean and Median of Child Count in the `infert` Dataset

**Task:** Calculate the mean and median of the `parity` column in the `infert` dataset.
<br>

**Steps:**

1. Load the `infert` dataset.
2. Calculate the mean and median of the `parity` column.

**Code Example:**

```{r}
# Load infert dataset
data(infert)

# Calculate mean and median of parity
mean_parity <- mean(infert$parity)
median_parity <- median(infert$parity)

mean_parity
median_parity
```
