```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(knitr.kable.NA = '')
library(knitr)
library(tidyr)
library(dplyr)
library(kableExtra)
```

```{=html}
<!--- When rendering to PDF, you may need to change "latex" to "latex" in 
kable commands--->
```

## Tidy Data Assignment {.unnumbered}

In this assignment, you will identify whether a given dataset is in tidy format. Each problem will present a dataset from a different background along with a brief description. Your task is to determine if the dataset is tidy. If it is not tidy, describe why and provide a tidy version of the dataset.

## Problems

### Problem 1: Weather Data

The following dataset contains weather data for three cities.

```{r echo=FALSE}
weather_data <- data.frame(
  City = c("New York", "Los Angeles", "Chicago"),
  Jan_Temp = c(30, 58, 25),
  Feb_Temp = c(32, 60, 28),
  Mar_Temp = c(45, 65, 40)
)

kable(weather_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin-right:auto;'")
```

<br>

### Problem 2: Student Grades

The following dataset contains grades for students in three subjects. <br>

```{r echo=FALSE}

grades_data <- data.frame(
  Student = c("Alice", "Bob", "Charlie"),
  Math = c(90, 85, 95),
  Science = c(88, 92, 89),
  History = c(84, 78, 91)
)

kable(grades_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

<br>

### Problem 3: Sales Data

The following dataset contains monthly sales data for different products.

```{r echo=FALSE}
sales_data <- data.frame(
  Product = c("A", "B", "C"),
  Jan_Sales = c(100, 150, 200),
  Feb_Sales = c(110, 160, 210),
  Mar_Sales = c(120, 170, 220)
)

kable(sales_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 4: Patient Health Data

The following dataset contains health data for patients.

```{r echo=FALSE}
health_data <- data.frame(
  Patient = c("John", "Jane", "Doe"),
  Height = c(170, 160, 180),
  Weight = c(70, 55, 80),
  Age = c(30, 25, 40)
)

kable(health_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 5: Financial Data

The following dataset contains quarterly financial data for companies.

```{r echo=FALSE}
financial_data <- data.frame(
  Company = c("X", "Y", "Z"),
  Q1_Revenue = c(1000, 2000, 3000),
  Q2_Revenue = c(1100, 2100, 3100),
  Q3_Revenue = c(1200, 2200, 3200)
)

kable(financial_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 6: Sports Statistics

The following dataset contains statistics for players in a sports team.

```{r echo=FALSE}
sports_data <- data.frame(
  Player = c("Player1", "Player2", "Player3"),
  Goals = c(5, 8, 7),
  Assists = c(3, 5, 4),
  Saves = c(2, 1, 3)
)

kable(sports_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 7: Movie Ratings

The following dataset contains ratings for movies by different critics.

```{r echo=FALSE}
ratings_data <- data.frame(
  Movie = c("Movie A", "Movie B", "Movie C"),
  Critic1 = c(4.5, 3.8, 4.7),
  Critic2 = c(4.0, 3.9, 4.8),
  Critic3 = c(4.7, 4.0, 4.9)
)

kable(ratings_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 8: Employee Salary Data

The following dataset contains salary data for employees in different departments.

```{r echo=FALSE}
salary_data <- data.frame(
  Employee = c("E1", "E2", "E3"),
  Dept1_Salary = c(50000, 55000, 60000),
  Dept2_Salary = c(52000, 57000, 62000),
  Dept3_Salary = c(54000, 59000, 64000)
)

kable(salary_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 9: Product Reviews

The following dataset contains reviews for products.

```{r echo=FALSE}
reviews_data <- data.frame(
  Product = c("Product1", "Product2", "Product3"),
  Review1 = c("Good", "Average", "Excellent"),
  Review2 = c("Very Good", "Good", "Very Good"),
  Review3 = c("Excellent", "Good", "Good")
)

kable(reviews_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 10: Course Enrollment Data

The following dataset contains enrollment data for courses.

```{r echo=FALSE}
enrollment_data <- data.frame(
  Course = c("Course1", "Course2", "Course3"),
  Semester1 = c(30, 25, 20),
  Semester2 = c(35, 28, 22),
  Semester3 = c(32, 26, 24)
)

kable(enrollment_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 11: Sales Data

The following table shows the monthly sales data for three products.

```{r, echo=FALSE}
sales2_data <- data.frame(
  Product = c("Product A", "Product B", "Product C"),
  January = c(100, 150, 200),
  February = c(120, 160, 220),
  March = c(130, 170, 230)
)

kable(sales2_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 12: Survey Data

The following table represents the results of a survey where respondents rated their satisfaction with three services.

```{r, echo=FALSE}
survey_data <- data.frame(
  Respondent = c("R1", "R2", "R3"),
  Service1_Satisfaction = c(5, 4, 3),
  Service2_Satisfaction = c(4, 3, 2),
  Service3_Satisfaction = c(3, 2, 1)
)

kable(survey_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 13: Weather Data

The table below shows the temperature readings at different times of the day for a week.

```{r, echo=FALSE}
weather2_data <- data.frame(
  Day = c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"),
  Morning = c(20, 21, 19, 22, 20),
  Noon = c(25, 26, 24, 27, 25),
  Evening = c(22, 23, 21, 24, 22)
)

kable(weather2_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 14: Exam Scores

The following table lists the scores of students in three subjects.

```{r, echo=FALSE}
exam_scores <- data.frame(
  Student = c("Student1", "Student2", "Student3"),
  Math = c(85, 90, 95),
  Science = c(88, 92, 96),
  History = c(80, 85, 90)
)

kable(exam_scores, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 15: Hospital Data

The table below shows the number of patients admitted to different wards of a hospital over three months.

```{r, echo=FALSE}
hospital_data <- data.frame(
  Ward = c("Ward A", "Ward B", "Ward C"),
  January = c(30, 25, 20),
  February = c(35, 30, 25),
  March = c(40, 35, 30)
)

kable(hospital_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 16: Marketing Data

The following table represents the results of a marketing campaign showing the number of leads generated from different channels.

```{r, echo=FALSE}
marketing_data <- data.frame(
  Channel = c("Email", "Social Media", "SEO"),
  Week1 = c(50, 60, 70),
  Week2 = c(55, 65, 75),
  Week3 = c(60, 70, 80)
)

kable(marketing_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 17: Fitness Data

The table below shows the workouts completed by three athletes over a week.

```{r, echo=FALSE}
fitness_data <- data.frame(
  Athlete = c("Athlete1", "Athlete2", "Athlete3"),
  Monday = c(30, 40, 50),
  Wednesday = c(35, 45, 55),
  Friday = c(40, 50, 60)
)

kable(fitness_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 18: Financial Data

The following table shows the quarterly profits for three companies.

```{r, echo=FALSE}
financial2_data <- data.frame(
  Company = c("Company A", "Company B", "Company C"),
  Q1 = c(10000, 15000, 20000),
  Q2 = c(12000, 16000, 22000),
  Q3 = c(13000, 17000, 23000),
  Q4 = c(14000, 18000, 24000)
)

kable(financial2_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 19: Attendance Data

The table below shows the attendance numbers for different events over three days.

```{r, echo=FALSE}
attendance_data <- data.frame(
  Event = c("Event A", "Event B", "Event C"),
  Day1 = c(100, 150, 200),
  Day2 = c(110, 160, 210),
  Day3 = c(120, 170, 220)
)

kable(attendance_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

### Problem 20: Production Data

The following table represents the production output of different products over three shifts.

```{r, echo=FALSE}
production_data <- data.frame(
  Product = c("Product X", "Product Y", "Product Z"),
  Shift1 = c(300, 400, 500),
  Shift2 = c(350, 450, 550),
  Shift3 = c(400, 500, 600)
)

kable(production_data, format = "latex", table.attr = "style='width:4in; margin-left:auto; margin.right:auto;'")  %>%
  kable_styling(position="center")
```

<br><br><br>
