---
title: "Web Scraping"
author: "Mike LeVan"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Web Scraping

Web scraping is the process of extracting information from websites. This 
can be done manually, but it is often done using a program or script. 
In this tutorial, we will use the `rvest` package in R to scrape data 
from a website.

### Installing the rvest package

To install the `rvest` package, use the following command:

```{r}

library(rvest)
```

### Reading HTML

To read HTML from a website, use the `read_html()` function. This function
takes a URL as an argument and returns an HTML document.



```{r}

url <- "https://www.basketball-reference.com/players/j/jamesle01.html"

lebron <- read_html(url)

lebron %>% 
  html_nodes("h1") %>% 
  html_text()

lebron %>%
  html_nodes("table") %>%
  html_table()

lebron %>%
  html_nodes("table") %>%
  html_table() %>%
  .[[1]]

lebron %>%
  html_nodes("table") %>%
  html_table() %>%
  .[[1]] %>%
  head()

```


